---
title: "CPP_analysis"
output:
  html_document: default
  pdf_document: default
---


```{r message=FALSE, include=FALSE}
install.packages("knitr")
install.packages("tidyverse") #1.2.1
install.packages("drc") #3.0.1
install.packages("stringr")
```

```{r message=FALSE}
library("drc") #3.0.1
library("knitr")
library("tidyverse") #1.2.1
library("stringr")
```

```{r datasheet}
all_data<-data.frame(read.csv("CPP_experiment_data_collated.csv"))
#Filter out empty/discarded oviposition papers 
blockAB<-dplyr::filter(all_data, TotalLarvaeEmbryos>0)
```

```{r yields}
#Calculate yield/fraction of initial count
blockAB$yield <- blockAB$TotalLarvaeEmbryos/blockAB$InitialEggs 
hist(blockAB$yield, main="Yield (final/initial count) per oviposition paper")
plot(blockAB$yield ~ blockAB$TotalLarvaeEmbryos)
```

```{r filter papers}
#removing papers with <60% or >100% initial count
blockAB <- filter(blockAB, yield<=1) %>% filter(yield>=0.6) 
```


```{r summarize}
DI_summary <- blockAB %>% 
  group_by(CageID) %>% 
  summarise(pop=first(Population), pp=mean(Photoperiod), lat=mean(Lat), block=first(Block), country=first(Country), h1=sum(H1Count), h2=sum(H2Count), emb=sum(EmbryoCount)) %>% 
  filter(h1+h2+emb>100)
DI_summary$total<-DI_summary$h1+DI_summary$h2+DI_summary$emb
DI_summary$DI<-DI_summary$emb/DI_summary$total
```



```{r totals_plot}
hist(DI_summary$total, main="Total number of viable eggs per cage")
  
```

```{r DI_plot}
#plot with DI at 8 h shown as TWO DIFFERENT points (corresponding to two cabinets)
ggplot(DI_summary, aes(x=pp, y=DI, group=pop)) + geom_point() + facet_wrap(~pop)
  
```

```{r incidence_lat}
DI_short<-filter(DI_summary, pp==8)
ggplot(DI_short, aes(x=lat, y=DI, color=country)) + geom_point() + theme_bw() +stat_smooth()
```

# Calculating CPP values

5-parameter generalized log-logistic model. Explanation on page 2: http://ftp.uni-bayreuth.de/math/statlib/R/CRAN/doc/vignettes/drc/drc.pdf
d and c are upper and lower limit, respectively
e is ED50
b is relative slope aroung e
f represents asymmetry in the model; f=1 if function is symmetrical 

```{r CPP_calc_2019}
#From Kevin Emerson

# the following fits the models of the given function using data from the
# data.frame d separately for the various treatments using a 5 parameter
# logistic model
d.model.fit <- drm(DI ~ pp, data = DI_summary, curveid = pop, fct = LL.5())
#plot(d.model.fit, log = "")

# output the parameters of the model fits:
summary(d.model.fit)

# estimate the 50% intercept
ED(d.model.fit, 50)
cpp_values<-data.frame(ED(d.model.fit, 50))

```

#Recalculating CPP values for 2008 photoperiod response curves

```{r CPP_calc_2008}
DI_summary_2008<-data.frame(read.csv("CPP_2008.csv"))
d.model.fit.2008 <- drm(DI ~ PP, data = DI_summary_2008, curveid = POP, fct = LL.5())
  
#plot(d.model.fit.2008, log = "")
# output the parameters of the model fits:
summary(d.model.fit)
# estimate the 50% intercept
ED(d.model.fit.2008, 50)
cpp_values_2008<-data.frame(ED(d.model.fit.2008, 50))
```
#Combine and format CPP value tables from 1988, 2008 and 2018

```{r collate_cpp_values}
cpp_2019<-data.frame(ED(d.model.fit, 50))
cpp_2019<-rownames_to_column(cpp_2019, var="pop") 
cpp_2019$pop<-str_sub(cpp_2019$pop,3,5)
popdata_2019<-DI_summary %>% group_by(pop) %>% 
  summarize(lat=first(lat), country=first(country))
cpp_2019<-right_join(cpp_2019, popdata_2019, by="pop")
#drop PBS from this list for now; CPP estimate doesn't make sense
#cpp_2019<-filter(cpp_2019, pop!="PBS")
cpp_2019$year<-rep("2019", nrow(cpp_2019))

cpp_2008<-data.frame(ED(d.model.fit.2008, 50)) %>% rownames_to_column(var="pop")
cpp_2008$pop<-str_sub(cpp_2008$pop,3,5)
popdata_2008<-DI_summary_2008 %>% group_by(POP) %>% 
  summarize(lat=first(LAT), country=first(COUNTRY), year=first(YEAR)) 
colnames(popdata_2008)[1]<-"pop"
cpp_2008<-right_join(cpp_2008, popdata_2008, by="pop")

cpp_1988<-data.frame(read.csv("CPP_1988.csv")) 
cpp_1988$pop<-na_if(cpp_1988$pop, "NA")
cpp_1988$Std..Error<-na_if(cpp_1988$Std..Error, "NA")


cpp_2<-rbind(cpp_1988, cpp_2008, cpp_2019) 

```

# CPP by latitude, country and year

```{r Japan_US_years}
ggplot(cpp_2, aes(x=lat, y=Estimate, fill=year, color=year)) + 
  scale_color_manual(values=c("white","gray","black")) + 
  scale_fill_manual(values=c("white","gray","black"))+ 
  geom_point(shape=21, color="black") + 
  geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), na.rm=TRUE) + 
  facet_grid(~country) + stat_smooth(method=lm, se=FALSE)
```
```{r }
ggplot(cpp_2, aes(x=lat, y=Estimate, color=country)) + 
  geom_point() + 
  geom_errorbar(aes(ymin=Estimate-Std..Error, ymax=Estimate+Std..Error), na.rm=TRUE) + 
  facet_grid(~year) + stat_smooth(method=lm, se=FALSE)
```

```{r incidence_lat_year}
DI_short_2008<-filter(DI_summary_2008, PP==8)
names(DI_short_2008)<-c("pop","country","lat","pp","year","DI")

DI_short_2018<-filter(DI_summary, pp==8) %>% group_by(pop) %>% 
  summarize(country=first(country), lat=first(lat), pp=first(pp), emb=sum(emb), total=sum(total))
DI_short_2018$DI<-DI_short_2018$emb/DI_short_2018$total
DI_short_2018<-DI_short_2018[,c(1:4,7)] 
DI_short_2018$year<-rep("2018",nrow(DI_short_2018))
DI_short_2018<-DI_short_2018[,c(1,2,3,4,6,5)]

DI_short_2<-rbind(DI_short_2008, DI_short_2018)

ggplot(DI_short_2, aes(x=lat, y=DI, color=year, label=pop)) + scale_color_manual(values=c("gray","black")) + 
  geom_point() + facet_grid(~country) 

```






